---
title: Using the command line
output: 
  html_document:
    toc: true
    toc_depth: 2
---

# Kernel Density Estimation

Most steps that are available from the GUI can also be called from the command line. 

```{r, message=FALSE}
library(rhr)
data(datSH)
str(datSH)
```
We see that `x` and `y` coordinates are saved in the second and third column. Next we will perform a KDE.

```{r}
kd1 <- rhrKDE(datSH[, 2:3])
```


The `kde` functions take three additional arguments: bandwidth, template raster and levels. 

## Bandwidth
The bandwidth can be provided as a vector of length one or two. Several methods to estimate bandwidth are available.

- `rhrHref`: estimates the reference bandwidth.
- `rhrHlscv`: estimates the least square cross validation bandwidth.
- `rhrHpi`: estimates the plug in the equation bandwidth. 
- `rhrHrefScaled`: estimates a scaled re fence bandwidth (as described in @Kie2013)

The default bandwidth is the reference bandwidth.

```{r}
## Calculating reference bandwidth:
(href <- rhrHref(datSH[, 2:3]))

## Or plugin the equation
(hpi <- rhrHpi(datSH[, 2:3]))

kdehref1 <- rhrKDE(datSH[, 2:3], h = href$h)
kdehref2 <- rhrKDE(datSH[, 2:3], h = rhrHref(datSH[, 2:3])$h)

## Are they the same
identical(rhrUD(kdehref1), rhrUD(kdehref2))
```

## Working with the result

Once we successfully estimated a KDE, there are several things we can do with. The first thing might be to plot it:

```{r}
plot(kdehref1)
```

The functions `rhrUD` and `rhrIsopleths` allows to retrieve the UD and isopleths respectively. 
```{r}
ud <- rhrUD(kdehref1)
class(ud)
```
The `ud` is a `RasterLayer` from the `raster` package. Similarly we can obtain the the isopleths at a predefined or specific level.

```{r}
## the default level of 95
iso1 <- rhrIsopleths(kdehref1)
plot(iso1)
class(iso1)

## other levels
iso2 <- rhrIsopleths(kdehref1, levels = seq(10, 90, 5))
plot(iso2)
```
The `rhrArea` functions returns the home-range area at a given isopleth. 

```{r}
## default level
rhrArea(kdehref1)

## other levels
rhrArea(kdehref1, levels = seq(10, 90, 5))
```

## Saving results

Isopleths can be saved various GIS formats, e.g., 

```{r, eval = FALSE}
raster::shapefile(iso1, "myShape.shp")

```


# Working with several animals/instances

Often it may be useful to work with more than one animal or instance (i.e., different tracking periods of the same animal) simultaneously. `rhr` does not provide direct support for this. However, it can be easily achieved using base R.

As an example, we will estimate monthly home ranges for the example red deer data set `datSH`. To do this we have to go through 3 steps:

1. Define a factor to split the data (e.g., species id, month of the year, study area).
2. Split the data with R's `split` function.
3. Apply some home range estimate.


```{r}
library(rhr)
library(lubridate)
data(datSH)

# 1. define a splitting variable, here I will use the months
datSH$month <- round_date(ymd(datSH$day), "month")

# 2. Split the data
dat2 <- split(datSH, datSH$month)

# 3. Apply home range estimates
hrs <- lapply(dat2, function(x) rhrKDE(x[, 2:3]))  # x is now the data for 1 month
```

`hrs` is now a list where each entry is the kernel density estimate for one month. To obtain the home range size (HRS) of the first month, we can use:

```{r}
rhrArea(hrs[[1]])
```

and with `sapply` we can easily obtain the home range size for all months.

```{r}
sapply(hrs, rhrArea)
```










