---
title: Working with tracks
output: 
  html_document:
    toc: true
    toc_depth: 2
---


# Create a track from a `*.csv` file

Suppose the relocation data is stored in a `*.csv` file, we want to load the data into R and create an `RhrTrackST` object. We will start with simulating some data.

## Generating data

This step can be omitted if you work with your own data.

```{r}
library(rhr)
set.seed(123321)

# Simulate a random walk
dat <- data.frame(rhrPoints(rhrRW()))

## Save the data as csv
write.csv(dat, file.path(tempdir(), "text.csv"))
```

## Reading the data and create a track object


```{r}
dat <- read.csv(file.path(tempdir(), "text.csv"))
head(dat)
```

In order to create `RhrTrackST` object, we need the coordinates as `SpatialPoints` and parsed time stamps. We will first create `SpatialPoints` for the coordinates.

```{r}
coords <- SpatialPoints(dat[, c("x", "y")])
```

Next we need to parse the time stamps, this is most easily done with the `lubridate` package.

```{r}
library(lubridate)
ts <- ymd_hms(dat$time)
```

Finally we can create a `RhrTrackST` object:

```{r}
track1 <- rhrTrack(coords, ts)
track1
```


# Working with tracks

## Plot

There is `plot` method for tracks.

```{r}
plot(track1)
```

Sometimes it my be desirable to map a certain attribute on the path. In the next example, we will use variable line with as a function of the step length. 

```{r}
trackSeg <- rhrSegments(track1, spatial = TRUE)
head(trackSeg)

library(scales)
plot(trackSeg, lwd = rescale(trackSeg$distance, c(0.01, 4)), 
     axes = TRUE)
```

Or we could color segments according to some path characteristic.

```{r}
cols <- colorRampPalette(c("blue", "red"))(length(trackSeg$distance))
cols <- cols[order(trackSeg$distance)]

plot(trackSeg, lwd = 2, col = cols,  axes = TRUE)
```


## Segments

We can extract the segments and stats there of with `rhrSegments`:

```{r}
seg1 <- rhrSegments(track1)
```

`seg1` is a `data.frame` with the start and end point of each segment, time and some basic statistics, such as 

- distance
- duration
- speed 
- direction.

We could for example plot the distance that animal moved over time. 

```{r}
plot(rhrTimes(track1)[-1], seg1$distance, type = "l", xlab = "time", ylab = "movement distance")
```


## Net Square displacement

The function `rhrNSD` allows the calculation of the the net square displacement.

```{r}
nsd <- rhrNSD(track1)
plot(nsd, type = "l")
```

# Tracks with properties

It is also possible to add some additional information to a track (e.g., behavioral state, landscape attributes). 


```{r}
dat <- read.csv(file.path(tempdir(), "text.csv"))
head(dat)
```

We proceed in the same way as before, but make use of a `SpatialPointsDataFrame`. 


```{r}
# simulate some env covariates
attr <- data.frame(mode = sample(c("feeding", "sleeping", "moving"), nrow(dat), TRUE), 
                   elevation = runif(100, 500, nrow(dat)))

coords <- SpatialPointsDataFrame(dat[, c("x", "y")], attr)
```

Next we need to parse the time stamps, this is most easily done with the `lubridate` package.

```{r}
ts <- ymd_hms(dat$time)
```

Finally we can create a `RhrTrackST` object:

```{r}
track1 <- rhrTrack(coords, ts)
track1
```

Lets get points and the segments back from the track:

```{r}
seg1 <- rhrSegments(track1, spatial = TRUE)
head(seg1)
```

The attributes are lost from `track1`. That is because `R` does not know what to do with it, because each segment has a start and end point where the attributes are measured. We can add easily add attributes to segments with by asking for the points with attributes:

```{r}
pts <- rhrPoints(track1)
head(pts)
```

Lets add the movement mode at the start (`mode_end`) and at the end  (`mode_start`) to each segment, we can do this with:

```{r}
seg1$mode_start <- pts$mode[-nrow(pts)]
seg1$mode_end <- pts$mode[-1]
```

Similarly we could also take the mean elevation of the start and end point:

```{r}
seg1$mean_ele <- (pts$elevation[-1] + pts$elevation[-nrow(pts)])/2
```

As shown before, we can use these attributes to create plots:

```{r}
plot(seg1, col = as.numeric(as.factor(seg1$mode_start)))
```

