---
title: Working with tracks
output: 
  html_document:
    toc: true
    toc_depth: 2
---


# Create a track from a `*.csv` file

Suppose the relocation data is stored in a `*.csv` file, we want to load the data into R and create an `RhrTrackST` object. We will start with simulating some data.

## Generating data

This step can be omitted if you work with your own data.

```{r}
library(rhr)
set.seed(123321)

# Simulate a random walk
dat <- data.frame(rhrPoints(rhrRW()))

## Save the data as csv
write.csv(dat, file.path(tempdir(), "text.csv"))
```

## Reading the data and create a track object


```{r}
dat <- read.csv(file.path(tempdir(), "text.csv"))
head(dat)
```

In order to create `RhrTrackST` object, we need the coordinates as `SpatialPoints` and parsed time stamps. We will first create `SpatialPoints` for the coordinates.

```{r}
coords <- SpatialPoints(dat[, c("x", "y")])
```

Next we need to parse the time stamps, this is most easily done with the `lubridate` package.

```{r}
library(lubridate)
ts <- ymd_hms(dat$time)
```

Finally we can create a `RhrTrackST` object:

```{r}
track1 <- rhrTrack(coords, ts)
track1
```

# Working tracks

## Plot

There is `plot` method for tracks.

```{r}
plot(track1)
```

## Segments

We can extract the segments and stat's there of with `rhrSegments`:

```{r}
seg1 <- rhrSegments(track1)
```

`seg1` is a `data.frame` with the start and end point of each segment, time and some basic statistics, such as 

- distance
- duration
- speed 
- direction.

We could for example plot the distance that animal moved over time. 

```{r}
plot(rhrTimes(track1)[-1], seg1$distance, type = "l", xlab = "time", ylab = "movement distance")
```



## Net Square displacement

The function `rhrNSD` allows the calcualtion of the the net square displacement.

```{r}
nsd <- rhrNSD(track1)
plot(nsd, type = "l")
```



