---
title: Installing `rhr`
output: 
  html_document:
    toc: false
    toc_depth: 1
---

# Requirements
In order to work with the `rhr` package, **R** needs to be installed on your system (the latest version of **R** is recommended, but at minimum version 3.0 is required). **R** can be [downloaded](http://www.r-project.org) for free. In addition to **R** a modern Internet Browser is required. All testing was done on Mozilla Firefox (version > 30), therefore we suggest to use Firefox, which can also be [downloaded](http://www.mozilla.org/) for free.

# Installation of the package
Currently the package is hosted on github and can be installed from there. We have also set up a repository hosting the `rhr` package. When you install the package for the first time, it is very likely that many additional packages are being installed. This is because the `rhr` package depends on several other **R**-packages, which in turn depend on further packages. Luckily **R**'s packaging system usually does a very good job and automatically installs all dependencies, so that we do not have to intervene here manually.


## Installation of `rhr` from an alternative repository
At the moment it is recommended to install the package from an alternative repository. This can be done by copying the following command and pasting it to **R**'s command line.
```{r, eval = FALSE}
  install.packages("rhr", repos=c("http://78.47.85.98/R",  
                                  "http://cran.rstudio.com/"), dep = TRUE)
```


## Installation for Mac users

First make sure all the depdencies are installed:

```{r, eval = FALSE}
avail <- installed.packages()[, 1]
needed <- c("grid", "maptools",  "raster", "adehabitatHR", "adehabitatLT", "ggplot2", "graphics", "gridExtra", "KernSmooth", "lubridate", "mixtools", "mvtnorm", "rgdal", "rgeos", "reshape2", "sp", "spacetime", "stringr", "trajectories", "zoo", "shiny", "knitr", "markdown", "testthat", "shinyBS", "brew", "xtable")

install <- needed[!needed %in% avail]

if (length(install) > 0) {
  install.packages(install)
}
```

Next install the `rhr` package:

```{r, eval = FALSE}
install.packages("http://78.47.85.98/rhr_1.2.900.tar.gz", repos = NULL, type = "source")
```

If it all worked, this should bring up the GUI:

```{r, eval = FALSE}
library(rhr)
rhrGUI()

```



