---
title: Site Fidelity
output: 
  html_document:
    toc: true
    toc_depth: 1
    mathjax: https://d3eoax9i5htok0.cloudfront.net/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML

---

```{r setup, include = FALSE}
library(rhr)
library(knitr)

knitr::opts_chunk$set(cache = TRUE)

```


# Background 

Implementation of site fidelity is based on suggested methods by [Spencer et al. 1990](http://www.jstor.org/stable/1937590). As a definition for the presence they use:

> Site fidelity exists, if the observed area an animal uses is smaller than the area used if an indiviudals movement was random ([Munger 1984](http://www.esajournals.org/doi/abs/10.2307/1938315)). 

They then assumed, that if an animal exhibits site fidelity, it posses an home range. To test for the existence of site fidelity, they used three metrics:

- A linearity index (LI)
- The mean squared distance from the center of activity (MSD)

In `rhr` the linearity index and the mean squared distance from the center of activity are implemented.

# Calculation

The test whether or not animal exhibits site fidelity the LI and MSD metric a compared to random trajectory. Randomly permuting step lengths and connected them with random angles. This is done $n = 100$ times and for each random trajectory LI and MSD are calculated to obtain a distribution of these values. Finally the values for LI and MSD are compared to the distribution LI and MSD for random trajectories.

```{r randomTraj, echo = FALSE}
data(datSH)
i1 <- rhr:::rhrBasePRW(datSH[1:500, 2], datSH[1:500, 3])

par(mfrow = c(1, 2))
plot(datSH[1:500, 2:3], type = "l", asp = 1, xlab = "x", ylab = "y", main = "Actual Path")
plot(i1, type  ="l", asp = 1, xlab = "x", ylab = "y", main = "Permuted Path")
par(mfrow = c(1, 1))
```


# Interpretation

`rhr` produces an histogram for LI and MSD and indicates where the critical threshold (red dashed lines) is located (by default using $\alpha = 0.05$) and the observed value (red solid line). 

```{r}
sf <- rhrSiteFidelity(datSH[1:500, 2:3], n = 500)
plot(sf)

```

If the solid red line is below the interval indicated by the red dashed lines, one can conclude that site fidelity is present. 

